An event-sourced and decentralized database with IPLD and materialized views

### 2019-09-27 draft.tex
updated el section
Signed-off-by: andrew <andrew@textile.io>

In multi-writer systems, conflicts arise as disparate peers end up producing conflicting state changes, or changes that end up out of sync. In order to proceed, there must be some way to deal with these conflicts. As mentioned previously (see Section \ref{sec:merkleclocks}), one way to avoid conflicts to some degree is to determine the total order of events in the shared log. In multi-writer systems, determining this causal order in all replicas is particularity challenging. In some cases (e.g, \texttt{ipfs-log} \cite{markroberthendersonOrbitDBFieldManual2019}), a Merkle-Clock can be used to induce ordering. However, this is only \emph{partially ordered}. That is, this approach cannot achieve a total order of events without implementing a data-layer conflict resolution strategy \cite{sanjuanMerkleCRDTs2019}:

@@

In many cases, the total order of an event sequence is meaningful, and this type of deterministic merge strategy is not sufficient. A git merge highlights one such case, in which additional information is sometimes needed in order to resolve a line conflict. Ledger based transactions present a similar challenge where the absolute order of transactions between multiple peers really does matter.

@@

% [!b] indicates that it will try to be on the bottom of the page
\begin{figure}[!b]
  \includegraphics[width=\linewidth]{figures/Event_Log.png}
  \caption{A single-writer Merkle-Clock (Event Log).}
  \label{fig:EventLog}
\end{figure}

Clearly, multi-writer systems with relaxed consistency \cite{brewerRobustDistributedSystems2000,gilbertBrewerConjectureFeasibility2002a} and imperfect information (e.g., caused by computational or network issues, both of which are common in p2p systems) will almost inevitably lead to situations where we need to decide how to reconcile a conflict. Some systems come with conflict resolution strategies ``built in'' (e.g., \texttt{ipfs-log} uses a built-in CRDT). However, this is not ideal, because it forces developers/users into a very specific usage pattern. It is particularly limiting in situations where developers might like to institute their own CRDT or domain-specific conflict resolution strategies.

>>>>

In multi-writer systems, conflicts arise as disparate peers end up producing conflicting state changes, or changes that end up out of sync. In order to proceed, there must be some way to deal with these conflicts. In some cases (e.g, \texttt{ipfs-log} \cite{markroberthendersonOrbitDBFieldManual2019}), a Merkle-Clock can be used to induce ordering. This approach cannot achieve a total order of events without implementing a data-layer conflict resolution strategy and therefor still requires, data-layer conflict resolution \cite{sanjuanMerkleCRDTs2019}:

@@

Therefore, solutions such as \texttt{ipfs-log} include a built-in CRDT to manage conflicts not resolved by the Merkle-Clock. This approach works for many cases, but the use of a deterministic merge strategy can be insufficient in cases where complicated data structures or long chains of missed events need to be reconciled. A git merge highlights one such example, in which a predetermined merge strategy can be used, but is not often the best choice in practice. Furthermore, a multi-writer log using a linked data format (e.g., a Merkle-Dag) in imperfect networking or storage environments can lead to suboptimal states where it may exceedingly costly (networking, storage, and computation) to regain consistency.

A promising approach is one that can leverage the benefits of both a Merkle-Clock for events from any one peer and a less constrained ordering mechanism to combine events from all peers. In this more ideal setup, developers can more freely institute their own CRDTs or domain-specific conflict resolution strategies. Additionally, it more naturally supports use-cases where all peers contributing to a dataset may not be interested in following or replicating the events of all other peers.

@@

% [!b] indicates that it will try to be on the bottom of the page
\begin{figure}[!t]
  \includegraphics[width=\linewidth]{figures/Event_Log.png}
  \caption{A single-writer Merkle-Clock (Event Log).}
  \label{fig:EventLog}
\end{figure}
